Testable normative theories

### 2020-06-01 simple.tex
Cleaned up various comments; clarified main claim

\subsection{Definition}
A normative theory is a tuple $(\allprops, ¬, \allargs, I, (\ileadstoe[i])_{i \in I})$, 
with ${\ileadstoe[i]} \subseteq \allhist × \Phi$, 
where $\allhist$ denotes all finite sequences of elements in $\allargs$ (including the empty sequence) and $\Phi = \allprops × \set{\text{possible}, \text{sure}}$.

@@

\commentOC{TODO also: in general, the model defines $\gargs[\phi]$ as a function of $\ileadstoe$? Start with $\gargs[\phi, 0]$, then observe all $\ar_1$ attacking those, and the model may define $\gargs[\phi, 1]$ depending on this observation, and so on. The condition is that one such argument be decisive.}

\subsection{Validity and conditions}
\begin{definition}[Validity]
	An empirical theory is \emph{valid} iff $\forall \phi \in {\gleadsto}(\allargs): [\phi \in \iPhi \land ¬\phi \notin \iPhi]$.
\end{definition}

\begin{remark}
	In fact, the model is tasked with finding a stable point after interrogating $i$. The process of trying counter-arguments $\ar_1, …$ is $q(i)$. When done, the model claims that $\ar_0 + \ar_2 + …$ is decisive. The model also claims that another process of interrogation does not lead to different answers. This is falsifiable.

	It follows that trying different $\ar_2, \ar'_2, …$ is also admissible.

	I need a (more) basic observable: a path $p$ is a sequence of arguments given to $i$, and I assume that whether $p \ileadsto[i] \phi$ is observable. But we can only assume that one path and its extensions is observable.

	When claiming that $(\ar_1, \ar_0) \ileadstost[i] \phi$, the model really claims that $\nexists p, i' \in q^{-1}(q(i)) \suchthat p, \ar_1, \ar_0 \ileadsto[i'] ¬\phi \lor p, \ar_0, \ar_1 \ileadsto[i'] ¬\phi$, where $i' \in q^{-1}(q(i))$ iff $q(i) = q(i')$.

	To say that $\phi \in \iPhi$, need to not depend on choice of $q$, in the following sense: if with $q$, $\phi$ seems stable, but with $q'$, $¬\phi$ seems stable, then $\phi$ is not proved as being in $\iPhi$. But if with $q$, $\phi$ seems stable, but with $q'$, nothing can be shown to be stable, then $\phi$ is proved (temporarily) as being in $\iPhi$. More precisely, after $q'$, the model must be able to stabilize $i$ with its $q$.
\end{remark}
I think I need the following. Define $\ar_1 \ibeatse \ar_0$ iff $\exists \hist \suchthat (\hist, \ar_0, \ar_1) \ileadstoe ¬\phi$. The model claims that if $\ar_1 \ibeatse \ar_0$ and $\ar_3 \ibeatse \ar_0 + \ar_2$, then $(\hist, \ar_0 + \ar_2 + \ar_4) \ileadstoe \phi$ (for some suitable $\hist$ that does not contain yet this arg or their c-a, or perhaps just claim that $\exists \hist \suchthat \hist \ileadstoe \phi$ with $\hist$ containing this arg).

Better: the model claims that if $\ar_1 \ileadstoe ¬\phi$ and if $\ar_0$ and $\ar_1$ are in $\hist$ and none of the anticipated attackers of $\ar_0 + \ar_2$ are, then $(\hist, \ar_0 + \ar_2) \ileadstoe \phi$ (or, at least, that the model has at least one available candidate).

>>>>

\subsection{Definition}
A normative theory is a tuple $(\allprops, ¬, \allargs, I, (\ileadstoe[i])_{i \in I})$, 
with ${\ileadstoe[i]} \subseteq \allhist × \Phi$, 
where $\allhist = \bigcup_{k \in \N} \allargs^{\intvl{1, k}}$ denotes all finite sequences of elements in $\allargs$ (including the empty sequence) and $\Phi = \allprops × \set{\text{possible}, \text{sure}}$.

@@

\subsection{Validity}
\begin{definition}[Validity]
	An empirical theory is \emph{valid} iff $\forall \phi \in {\gleadsto}(\allargs): [\phi \in \iPhi \land ¬\phi \notin \iPhi]$.
\end{definition}

\subsection{Conditions with history}
\begin{remark}
	In fact, the model is tasked with finding a stable point after interrogating $i$. The process of trying counter-arguments $\ar_1, …$ is $q(i)$. When done, the model claims that $\ar_0 + \ar_2 + …$ is decisive. The model also claims that another process of interrogation does not lead to different answers. This is falsifiable.
	When claiming that $(\ar_1, \ar_0) \ileadstost[i] \phi$, the model really claims that $\nexists p, i' \in q^{-1}(q(i)) \suchthat p, \ar_1, \ar_0 \ileadsto[i'] ¬\phi \lor p, \ar_0, \ar_1 \ileadsto[i'] ¬\phi$, where $i' \in q^{-1}(q(i))$ iff $q(i) = q(i')$.

	To say that $\phi \in \iPhi$, need to not depend on choice of $q$, in the following sense: if with $q$, $\phi$ seems stable, but with $q'$, $¬\phi$ seems stable, then $\phi$ is not proved as being in $\iPhi$. But if with $q$, $\phi$ seems stable, but with $q'$, nothing can be shown to be stable, then $\phi$ is proved (temporarily) as being in $\iPhi$. More precisely, after $q'$, the model must be able to stabilize $i$ with its $q$.
\end{remark}

Given $\ar_0 \in \allargs$ and $\args_2 \subseteq \allargs$, define the set $\args_1 \subseteq \allargs$ of arguments that $\args_2$ protects $\ar_0$ from as $S_1 = [f^{-1}(\args_2)]^{-1}(\ar_0) = \set{\ar_1 \in \allargs \suchthat f(\ar_1, \ar_0) \cap \args_2 ≠ \emptyset}$. 
Define $g(\ar_0) \subseteq \allargs × \allargs$ as $g(\ar_0) = \set{(\ar_2, \ar_1) \suchthat \ar_2 \in f(\ar_1, \ar_0)} = \bigcup_{\ar_1 \in \allargs} (f(\ar_1, \ar_0) × \set{\ar_1})$. The set $g(\ar_0)(\allargs)$ are the planned attacks on $\ar_0$ and the set $g(\ar_0)^{-1}(\ar_1)$ are the planned protections of $\ar_0$ against $\ar_1$. 
Then, $\args_1 = g(\ar_0)(\args_2)$.

$\ar_0$ defends $\phi$ under protection of $S_2$ iff $\forall \hist \in \allhist: [\args_1 \cap \hist = \emptyset] ⇒ (\hist, \ar_0) \ileadstoe \phi$, where $\args_1 = g(\ar_0)(\args_2)$.

\begin{definition}[Claim]
	$\forall i \in I, \exists k \in 2\N$ and a finite sequence $(\args_j)_{j \in 2\N, j ≤ k} \suchthat \args_0 \subseteq {\gleadstoinv}(\phi)$ and $\forall j \in 2\N, j ≤ k, \forall \ar_j \in S_j: \ar_j$ defends $\phi$ under protection of $S_{j + 2}$ (defining $S_{k + 2} = \emptyset$) and $S_j \subseteq \cup_{\ar_{j - 2} \in \args_{j - 2}, \ar_{j - 1} \in \allargs} f(\ar_{j - 1}, \ar_{j - 2})$.
\end{definition}
Note that requiring that $\args_0$ be a singleton does not modify the condition, in the sense that any theory satisfying the condition also satisfies its restriction.

\begin{remark}
	The claim must have the form: some structure is adequate for all individuals. Here, the structure is given by $f$. The claim should claim that looking at the attacks planned by $f$ is enough; and looking at the defenses planned by $f$ is enough.

	This may be adequate as if $\ar_0$ seems suitable but some unplanned attacker $\ar_1$ attacks it, then use $\ar_1 \in \hist$ and observe that $¬(\hist, \ar_0) \ileadstoe \phi$ to deny the claim.

	Perhaps this could be phrased as follows. Say that $\ar_2$ is replaceable by $\args_2$ iff what $\ar_2$ protects is either protected by $\args_2$ or replaceable (by what?). Say that a set $\args$ is sufficient iff its complement is replaceable by $\args$. Then $g$ claims (among others) that $g(\ar_0)^{-1}(\ar_1)$ is sufficient.

	The claim can (hopefully) be given the form: $\forall i \suchthat …, \ar$ decisive. Given $i$, either $\ar_0$ is decisive, or $\ar_1$ attacks it but then $\ar_2$ is decisive, …
\end{remark}

\subsection{Conditions without history}
To verify or falsify these conditions, whether $(\ar_1, \ar_0) \ileadstoe \phi$ must be observable, but one may suspect that once $i$ has seen some $\ar$, then no more $(\ar_1, \ar_0) \ileadstoe \phi$ but that this does not imply that this would hold without $\ar$ for $i$. The general falsifiable claim is not clear.
